# Data preperation {#intro}


##Data
We can enter CSV data in R using read.csv function or read_csv function from readr package. However, for now we will use data that already exist in the datasets packages that we downloaded

Lets load the library 'datsets'
```{r include=FALSE}
library(datasets)
library(tidyverse)


```


Lets look at the chickweight data 

```{r include=TRUE}
head(ChickWeight)
str(ChickWeight)
```

##Data manipulation with dplyr and tidyverse

Let us create a column that calculates number of chicks following each type of diet

Lets look at the chickweight data 

```{r include=TRUE}
ChickWeight2<-ChickWeight %>% 
                 count(Diet)
   
head(ChickWeight2)
```

Let's look at mean weight of chick in each diet category

```{r include=TRUE}
ChickWeight3<-ChickWeight %>% 
              group_by(Diet) %>% 
              mutate(mean.wt= mean(weight)) # muatate can be used to create new columns
   
head(ChickWeight3)
```
Let's arrange the data in wide form, ggplot often likes the data to be in long form, that is stacked in rows. 
For example, here we have Diet in rows.

Let's look at some data in wide form
```{r include=TRUE}
head(relig_income)
```
Let's arrange so that income is stacked in rows
```{r include=TRUE}
relig_income2<-relig_income %>% pivot_longer(!religion, names_to="income",values_to="count")
head(relig_income2)
```
Re-do the wide 
```{r include=TRUE}
relig_income.back<-relig_income2 %>% pivot_wider( names_from="income",values_from="count")
head(relig_income.back)
```

