--- 
title: "Plotting in R using ggplot"
author: "Vratika Chaudhary"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a practical example of creating impactful visualizations for
  presenting results of scientific data analysis. Examples mainly use ggplot2 package
  in R."
---

# Prerequisites

R programming language
Installed packages:

  
  

```{r eval=FALSE}
pkg<-c("ggplot2","ggpubr","auk","datasets","ggpubr","ggcorrplot")
install.packages(pkg)

```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'auk', 'knitr', 'rmarkdown' ,'tidyverse','ggplot','datasets','ggpubr','ggcorrplot'
), 'packages.bib')
```


<!--chapter:end:index.Rmd-->

# Data preperation {#intro}


##Data
We can enter CSV data in R using read.csv function or read_csv function from readr package. However, for now we will use data that already exist in the datasets packages that we downloaded

Lets load the library 'datsets'
```{r include=FALSE}
library(datasets)
library(tidyverse)


```


Lets look at the chickweight data 

```{r include=TRUE}
head(ChickWeight)
str(ChickWeight)
```

##Data manipulation with dplyr and tidyverse

Let us create a column that calculates number of chicks following each type of diet

Lets look at the chickweight data 

```{r include=TRUE}
ChickWeight2<-ChickWeight %>% 
                 count(Diet)
   
head(ChickWeight2)
```

Let's look at mean weight of chick in each diet category

```{r include=TRUE}
ChickWeight3<-ChickWeight %>% 
              group_by(Diet) %>% 
              mutate(mean.wt= mean(weight)) # muatate can be used to create new columns
   
head(ChickWeight3)
```
Let's arrange the data in wide form, ggplot often likes the data to be in long form, that is stacked in rows. 
For example, here we have Diet in rows.

Let's look at some data in wide form
```{r include=TRUE}
head(relig_income)
```
Let's arrange so that income is stacked in rows
```{r include=TRUE}
relig_income2<-relig_income %>% pivot_longer(!religion, names_to="income",values_to="count")
head(relig_income2)
```
Re-do the wide 
```{r include=TRUE}
relig_income.back<-relig_income2 %>% pivot_wider( names_from="income",values_from="count")
head(relig_income.back)
```


<!--chapter:end:01-intro.Rmd-->

# Introduction to ggplot2

The package ggplot2 is extremely functional and useful. It created graphs in layers and therefore is very intutive. It adds layers on top of each other 

## Creating the building blocks of the graph

ggplot(data = dataname, mapping= aes(x=xaxis, y= yaxis))

ggplot function from ggplot2 package requires two arguments:
  data is the object name you gave to your data sets
  Mapping is the specifications of the variables we want to visualize and are places withing the  'aes' function short for aesthetics.

Let's bring back the chick data and let's look how chicks have grown over time

```{r nice-fig, fig.cap='Building block', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)
par(mar = c(4, 4, .1, .1))
ggplot(data=ChickWeight, mapping=aes(x=Time, y=weight))
```
We see a box, and we see the correct axis but we don't see anything else- why?

## Adding geom_objects
Because yes we told ggplot what variables we want to plot, but we haven't told it what type of plot we want. 
We can do this by adding a 'geom_object'. For this purpose, let's make a line plot. We can do so by adding a 'geom_line' object to the previous layer. We join the two layers by a '+' sign.

```{r, fig.cap='Line plot', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, mapping= aes(x=Time, y=weight))+
  geom_line()
```

Hey! We got some lines, but it looks weird!!!
That is becasue we plotted all the data together, we had 48 chicks but we plotted all of them in the same plot. So let's make them distinct. We can simply do so by giving a distinct line to each chick by adding group

```{r, fig.cap='Line plot with colors', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, aes(x=Time, y=weight, group=Chick))+
  geom_line()
```

## Making the plot visually more appealing 

Let's do the following:
1. Increase the size of the lines
2. Lighten the background using the theme function in ggplot2
3. Increase the size of axis text and titles
4. Change y axis label
```{r, fig.cap='Better looking line plot with groups', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, aes(x=Time, y=weight, group=Chick),size=2)+
  geom_line()+
  theme_minimal(base_size = 20)+
  ylab("Weight")
```

## Let's make the plot  colorful

```{r, fig.cap='Better looking line plot with colors', out.width='80%', fig.asp=.75, fig.align='center'}


ggplot(ChickWeight, aes(x=Time, y=weight, color=Chick),size=2)+
  geom_line()+
  theme_minimal(base_size = 20)+
  ylab("Weight")
```

## Change label fonts, color, angle etc

```{r, fig.cap='change label font', out.width='80%', fig.asp=.75, fig.align='center'}


ggplot(ChickWeight, aes(x=Time, y=weight, color=Chick),size=2)+
  geom_line()+
  theme_minimal(base_size = 20)+
  ylab("Weight")+theme(axis.title.x = element_text( angle=40,vjust=0.2,color='blue',family="Comic Sans MS"))+
  theme(axis.title.y =  element_text( color='purple',family="serif"))
```


## You can also create and save a theme for entire project

```{r, fig.cap='change label font', out.width='80%', fig.asp=.75, fig.align='center'}
theme_set<-function(){ theme_bw(base_size = 20)+
    theme(axis.text.x = element_text(angle=35,vjust=0.8,color='green'))+
    theme(legend.title = element_blank())}


ggplot(ChickWeight, aes(x=Time, y=weight, color=Chick),size=2)+
  geom_line()+theme_set()
```

## Changing colors manually and add error bars

Lets plot the mean chick weight over time for each type of diet

```{r, fig.cap='change label font', out.width='80%', fig.asp=.75, fig.align='center'}

ChickWeight %>% group_by(Diet,Time) %>% mutate(mean.wt=mean(weight), 
                  se.weight= (sd(weight)) / sqrt((length(weight)))) %>% 
                 ggplot() + 
              geom_line(aes(x=Time, y=mean.wt, color=Diet,lty=Diet),size=2)+
            geom_ribbon(aes(x=Time, ymax=mean.wt+se.weight,ymin= mean.wt-se.weight,group=Diet,fill=Diet),size=2,alpha=0.2)+
              scale_color_manual(values=c("cyan1","cyan4","green","orange"))+
            scale_fill_manual(values=c("cyan1","cyan4","green","orange"))+
              ggpubr::theme_pubr(base_size = 20)+
                ylab("Mean weight")
                     
```


<!--chapter:end:02-literature.Rmd-->

# Plotting using one variable at a time

## Histograms- plots the distribution of a numerical variable

Note the customizations of the histogram:
1. we colored the bars pretty
2. 'alpha' adds transparency to the object, useful when you have overlapping objects, goes from 0 (transparent) to 1 (opaque)

```{r, fig.cap='Histogram', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, aes(weight))+
  geom_histogram(fill='cyan4',color='black',alpha=0.5)+
  theme_minimal(base_size = 20)+
  ylab("Frequency")+ xlab("Weight (g)")
```

Improving 'binwidth' value, it is the number of categories data is divided into and usually defaults to 30, but we can go up and down, let's go to 50 here
```{r, fig.cap='Histogram', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, aes(weight))+
  geom_histogram(fill='cyan4',color='black',alpha=0.5, binwidth = 50)+
  theme_minimal(base_size = 20)+
  ylab("Frequency")+ xlab("Weight (g)")
```


## Bar plots 

Plot the distribution of a categorical variable

Here we will plot number of chicks following a particular diet.
```{r, fig.cap='Bar plot', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(ChickWeight, aes(x= Diet))+
  geom_bar(color='orange', fill='lavender')+
  theme_minimal(base_size = 20)+
  ylab("Count")+ xlab("Diet")
```
## Ordering based on number of counts from lowest to highest

Let's bring the magic of dplyer to do so

So we reorganize the data to calcualte number of chicks on each diet type



```{r}

library(tidyverse)
reord.chick<-ChickWeight %>% 
 count(Diet) %>% arrange(n)
reord.chick


```

Here we plot using geom_col function where heights represents the value of the data and requires y aesthetics. We can use geom_text function to add the label to each column and use vjust to move the labels up and down

```{r, fig.cap='Bar plot with labels', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(reord.chick, aes(x= reorder(Diet,n), y=n))+
  geom_col(color='orange', fill='lavender')+
  geom_text(aes(label=n), vjust=-0.7)+
  theme_minimal(base_size = 20)+
  ylab("Count")+ xlab("Diet")

```
## Adding percentages

Here we also get more adventerous with data wrangling and create new columns and plot them within the same pipe
```{r, fig.cap='Bar plot with percent labels', out.width='80%', fig.asp=.75, fig.align='center'}

ChickWeight %>% 
  count(Diet) %>% arrange(n) %>% 
mutate(percent = n / sum(n),
         percentlabel = paste0(round(percent*100), "%")) %>% 

ggplot( aes(x= reorder(Diet,percent), y=percent))+
  geom_col(color='orange', fill='lavender')+
  geom_text(aes(label=percentlabel), vjust=-0.7)+
  theme_minimal(base_size = 20)+
  ylab("Count")+ xlab("Diet")

```

## pie chart
```{r, fig.cap='Pie chart with labels ', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggplot2)

ggplot(reord.chick, aes(x= reorder(Diet,n), y=n,fill=Diet))+
  geom_col(width = 1, 
           stat = "identity", 
           color = "black") +
  geom_text(aes(label=n), vjust=-0.7)+
  theme_minimal(base_size = 20)+
  coord_polar("y", 
              start = 0,direction = -1 
              ) +
  theme_void() +
  ylab("Count")+ xlab("Diet")

```


<!--chapter:end:03-method.Rmd-->

# Multivariate plotting
Now we will plot using multiple variables

Let's use diamond dataset from tidyr
```{r}
data("diamonds")
head(diamonds)
```

## Bar plot with categories, plot depth by cut


```{r, fig.cap='Bar plot with grouping', out.width='100%', fig.asp=.75, fig.align='center'}




ggplot(diamonds)+
 geom_bar(aes(x= cut, fill=color))+
  theme_minimal(base_size = 20)+
  ylab("Count")+ xlab("Cut")

```


## Bar plot with categories, side by side


```{r, fig.cap='Bar plot with grouping', out.width='100%', fig.asp=.75, fig.align='center'}


ggplot(diamonds)+
 geom_bar(aes(x= cut, fill=color),position = position_dodge())+
  theme_minimal(base_size = 20)+
  ylab("Count")+ xlab("Cut")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) # when the axis text are overlapping, this can help

```


## Segemented bar plot, appealing viz

```{r, fig.cap='Bar plot with segments', out.width='75%', fig.asp=.75, fig.align='center'}


diamonds %>%
  group_by(cut, color) %>%
  summarize(n = n()) %>% 
  mutate(prct = n/sum(n),
         label = scales::percent(prct)) %>% 
         ggplot()+
          geom_col(aes(x=cut,y=prct,fill=color),position='fill')+
          geom_text(aes(x=cut,y=prct,label = label), 
            size = 3, color='white',
            position = position_stack(vjust = 0.5)) +
             theme_minimal(base_size = 20)+
  ylab("Percentage")+ xlab("Cut")


```




## Scatter plot

```{r, fig.cap='Scatter plot with grouping', out.width='80%', fig.asp=.75, fig.align='center'}

ggplot(diamonds)+
geom_point(aes(x=carat,y=price,color=color))+
  theme_bw()+
  xlab("Carat")+ ylab("Price")


```

Let's see if the Carat is related to price
```{r, fig.cap='Scatter plot with grouping and smooth line', out.width='80%', fig.asp=.75, fig.align='center'}

         ggplot(diamonds)+
          geom_point(aes(x=carat,y=price,color=color))+
  geom_smooth(aes(x=carat,y=price,color=color),method='lm')+
  theme_bw()+
  ylab("Carat")+ xlab("Price")


```
## Grouping using facets
```{r, fig.cap='Scatter plot with facets and smooth line', out.width='80%', fig.asp=.75, fig.align='center'}

         ggplot(diamonds)+
          geom_point(aes(x=carat,y=price))+
  geom_smooth(aes(x=carat,y=price),method='lm')+
  facet_wrap(~color)+
  theme_bw()+
  ylab("Carat")+ xlab("Price")


```

## Change facet labels
```{r, fig.cap='Scatter plot with facets and different labels', out.width='80%', fig.asp=.75, fig.align='center'}

diamonds2<-diamonds
levels(diamonds2$color)
diamonds2$color<- factor(diamonds2$color, levels =c("D" ,"E" ,"F" ,"G", "H", "I", "J"), 
                         labels=c("Red","Blue","Orange","Pink","Indigo","Jade","Orange"))
         ggplot(diamonds2)+
          geom_point(aes(x=carat,y=price))+
  geom_smooth(aes(x=carat,y=price),method='lm')+
  facet_wrap(~color)+
  theme_bw()+
  ylab("Carat")+ xlab("Price")


```

<!--chapter:end:04-application.Rmd-->

# Publication style figures and saving

'ggpubr' package is wondeful to create publication quality figures

```{r, fig.cap='Scatter plot with facets pub quality', out.width='80%', fig.asp=.75, fig.align='center'}

diamonds2<-diamonds
levels(diamonds2$color)
diamonds2$color<- factor(diamonds2$color, levels =c("D" ,"E" ,"F" ,"G", "H", "I", "J"), 
                         labels=c("Red","Blue","Orange","Pink","Indigo","Jade","Orange"))
         ggplot(diamonds2)+
          geom_point(aes(x=carat,y=price))+
  geom_smooth(aes(x=carat,y=price),method='lm')+
  facet_wrap(~color)+
  ggpubr::theme_pubr(base_size=22)+
           
  ylab("Carat")+ xlab("Price")


```


##Correlation plot
```{r}

diamonds2<-diamonds %>% select(depth,table, price,carat)

# calulate the correlations
c <- cor(diamonds2, use="complete.obs")



```

```{r, fig.cap='Correlation plot', out.width='80%', fig.asp=.75, fig.align='center'}
library(ggcorrplot)
ggcorrplot(c,lab=T, color=c("green","black","orange"))

```

<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

