--- 
title: "Plotting in R using ggplot"
author: "Vratika Chaudhary"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a practical example of creating impactful visualizations for
  presenting results of scientific data analysis. Examples mainly use ggplot2 package
  in R."
---

# Prerequisites

R programming language
Installed packages:

  
  

```{r eval=FALSE}
pkg<-c("ggplot2","ggpubr","auk","datasets","ggpubr","ggcorrplot")
install.packages(pkg)

```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'auk', 'knitr', 'rmarkdown' ,'tidyverse','ggplot','datasets','ggpubr','ggcorrplot'
), 'packages.bib')
```

<!--chapter:end:index.Rmd-->

# Data preperation {#intro}



We can enter CSV data in R using read.csv function or read_csv function from readr package. However, for now we will use data that already exist in the datasets packages that we downloaded

Lets load the library 'datsets'
```{r include=TRUE}
library(datasets)


```


Lets look at the chickeweight data 

```{r include=TRUE}
head(ChickWeight)

```

<!--chapter:end:01-intro.Rmd-->

# Literature

Here is a review of existing methods.

<!--chapter:end:02-literature.Rmd-->

# Methods

We describe our methods in this chapter.

<!--chapter:end:03-method.Rmd-->

# Applications

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:04-application.Rmd-->

# Final Words

We have finished a nice book.

<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

